IFA Comparison of Noncognitive Skills \\ BCS and MCS

### 2018-02-08 codes/R/cohorts_nc.Rnw
restructure file to put most stuff in appendix, add temporary notes

<<MAIN,cache=FALSE,echo=FALSE,results='hide',message=FALSE,warning=FALSE>>=
rm(list=ls())         # Clear all objects from memory 
  opts_chunk$set(concordance=FALSE,
                 echo=FALSE,
                 tidy=FALSE,
                 autodep=TRUE,
                 cache=T, # REPLACE LATER
                 message=FALSE, 
                 warning=FALSE,
                 size='scriptsize',
                 results='hide',
                 comment=NA, 
                 fig.width=4,
                 fig.height=4, 
                 out.width='.5\\textwidth',
                 fig.show='hold',
                 tikzDefaultEngine='pdftex',
                 tikzDocumentDeclaration = "\\documentclass[10pt]{article}",
                 tikzMetricPackages = c("\\usepackage[utf8]{inputenc}","\\usepackage[T1]{fontenc}", "\\usetikzlibrary{calc}", "\\usepackage{amssymb}")
                 )

@@

<<READ, cache=FALSE >>=
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/1_load.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/2_efa.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_mifa_invariance.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_mifa_final.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/4_scores.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/5_graphs.R")

@@

@@

\begin{table}[ht]
\centering
\small
\caption{Fit indices}\label{tab:fit}
<<FA_FIT_RES, dependson="FA_FIT", results='asis'>>=
require(xtable)

table <- rbind(afitab[[1]],afitab[[2]],afitab[[3]],afitab[[4]])

print(xtable(table,
                    align="L{0cm}L{.18\\textwidth}C{.05\\textwidth}C{.07\\textwidth}C{.07\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}",
                    digits = 4
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,4,8,12)),
                              command = c(
                              # header
            "Model & Num. params & $\\chi^2$ & RMSEA & MFI & CFI & G-hat & \\textDelta MFI & \\textDelta CFI & \\textDelta G-hat \\\\ \\midrule
                      \\newline \\textbf{No gender split} \\\\",
                      "\\newline \\textbf{Males only} \\\\ \\midrule ",
                      "\\newline \\textbf{Females only} \\\\ \\midrule ",
                      "\\newline \\textbf{Joint (4 groups)} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

@@

\begin{table}[ht]
\caption{Estimated loadings and thresholds} \label{tab:allpars}
\resizebox{\textwidth}{!}{
<<RES_FA_PAR, dependson="FA_FINAL_PARS", results='asis'>>=
require(xtable)

table <- rbind(allpar[[1]],allpar[[2]],allpar[[3]])

@@

                                            & $\\bm{\\lambda}$
                                            & $\\bm{\\tau}_1$
                                            & $\\bm{\\tau}_2$
                                            & $\\bm{\\Var(\\varepsilon)}$ \\\\
                                            Measure & BCS = MCS
                                            & BCS = MCS
                                            & BCS = MCS
                                            & MCS (BCS=1)
                                            & Factor
                                            \\\\ \\midrule
                      \\newline \\textbf{No gender split} \\\\",
                      # additional separators
                      "\\newline \\textbf{Males only} \\\\ \\midrule ",
                      "\\newline \\textbf{Females only} \\\\ \\midrule " 
                      )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabularx", width = "\\textwidth",
             floating = FALSE
)

@@

\begin{table}[ht]
\caption{Estimated loadings and thresholds -- 4-group model} \label{tab:allpars2}
\resizebox{\textwidth}{!}{
<<RES_FA_PAR2, dependson="FA_FINAL_PARS", results='asis'>>=
require(xtable)

table <- allpar[[4]]

@@

print.xtable(xtable(table,
                    # align = "ll|cccc|cccc"
                    align="L{0cm}L{1.5cm}C{2cm}C{1.5cm}C{1.5cm}C{1.8cm}C{1.8cm}C{1.8cm}C{1cm}",
                    digits = 3
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
                                          " & \\textbf{Loadings}
                                            & \\multicolumn{2}{c}{\\textbf{Thresholds}}
                                            & \\multicolumn{3}{c}{\\textbf{Variances}}
                                            & \\\\ \\cmidrule(lr){2-2} \\cmidrule(lr){3-4} \\cmidrule(lr){5-7}

@@

                                            Measure 
                                            & All groups
                                            & All groups
                                            & All groups
                                            & BCS Males
                                            & MCS Females
                                            & MCS Males
                                            & Factor
                                            \\\\"
                      )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabularx", width = "\\textwidth",
             floating = FALSE
)

@@

\begin{table}[ht]
\caption{Estimated latent variable means and covariances} \label{tab:lvpars}
\centering
<<RES_FA_LV, dependson="FA_FINAL_PARS", results='asis'>>=
require(xtable)

table <- rbind(lvpars[[1]],lvpars[[2]],lvpars[[3]],lvpars[[4]],lvpars[[5]])

>>>>

<<MAIN,cache=FALSE,echo=FALSE,results='hide',message=FALSE,warning=FALSE>>=
rm(list=ls())         # Clear all objects from memory 
  opts_chunk$set(concordance=FALSE,
                 echo=FALSE,
                 tidy=FALSE,
                 autodep=TRUE,
                 cache=T, # REPLACE LATER
                 message=FALSE, 
                 warning=FALSE,
                 size='scriptsize',
                 results='hide',
                 comment=NA, 
                 fig.width=4,
                 fig.height=4, 
                 out.width='\\textwidth',
                 fig.show='hold',
                 tikzDefaultEngine='pdftex',
                 tikzDocumentDeclaration = "\\documentclass[10pt]{article}",
                 tikzMetricPackages = c("\\usepackage[utf8]{inputenc}","\\usepackage[T1]{fontenc}", "\\usetikzlibrary{calc}", "\\usepackage{amssymb}")
                 )

@@

<<READ, cache=FALSE >>=
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/1_load.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/2_efa.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_mifa_invariance.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/3_mifa_final.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/4_scores.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/4_scores_compare.R")
read_chunk("/Users/giacomomason/Documents/Projects/CohortStudies/codes/R/5_graphs.R")

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\subsection{Inequality in noncognitive ability}
The model can be used to investigate how inequality of noncognitive skills has changed between the two cohorts.

\textbf{REVIEW} We can derive scores for $\theta^{EXT}$ and $\theta^{INT}$ directly, and plot them against socioeconomic measures observed in the data. Our main socioeconomic indicator is a measure of weekly total family net income in the prices of the interview year.\footnote{This measure is harmonised across cohorts and derived as part of CLOSER, which is funded by the ESRC.} Using the estimated parameters for the model with invariant thresholds and loadings (but non-invariant intercepts), I compute the scores for each individual in the sample (using the Empirical Bayes modal approach). In \autoref{fig:boxscores}, we can see that the gradient in noncognitive skills along family income quintiles is more pronounced in the MCS cohort. 
%In particular, a male five-year old at the top quintile of the income distribution in 1975 has a better score by \Sexpr{regcoeffs[1,3]} in the externalising dimension and \Sexpr{regcoeffs[1,4]} in the internalising dimension, compared to a child at the bottom quintile. In 2005, the differences amount to \Sexpr{regcoeffs[3,3]} and \Sexpr{regcoeffs[3,4]}, respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

List of estimated models:
\bnm
\item no gender split, no age adjustment (2 groups, BCS MCS)
\item no gender split, age adjustment (2 groups, BCS MCS)
\item males only, no age adjustment (2 groups, BCS MCS)
\item females only, no age adjustment (2 groups, BCS MCS)
\item males only, age adjustment (2 groups, BCS MCS)
\item females only, age adjustment (2 groups, BCS MCS)
\item separate gender groups, no age adjustment (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item separate gender groups, age adjustment (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item separate gender groups, overlapping ages, no age adjustment (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item separate gender groups, age adjustment constrained to be the same across ages (4 groups, BCS.M BCS.F MCS.M MCS.F)
\item MIMIC model with age as regressor (4 groups, BCS.M BCS.F MCS.M MCS.F)
\enm

Notes: 
\bi
\item Only the models with age adjustment seem to have invariant intercepts (see Appendix \autoref{sec:app_fit}). 
\item But the latent means of these models don't make sense with the "raw" scores we have (see Appendix \autoref{sec:app_meancov})
\item Also, a model on the subset of children that have very similar ages delivers results that are very similar to the age-unadjusted models, and different from the age-adjusted ones. So the models with age adjustment seem to be doing something else than adjusting for age, and we don't understand what this something else is...
\item The model with free intercepts (only restricting thresholds and loadings) delivers differences in scores by group, even if the means of the latent factors are constrained to be equal (see below)
\item The models with free and constrained intercepts deliver very similar scores (.98 correlation), and they also give sensible and comparable gradients with income
\ei

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<FA_SCORES_COMPARE, dependson='FA_SCORES'>>=
# assemble and compare scores across different models

When adjusting for age, mean scores do not agree with raw scores
<<FA_SCORES_MEANS, dependson='FA_SCORES', echo=T, results='markup'>>=
# means
means.ext
means.int
@

Scores that do not adjust for age are very highly correlated, regardless of whether they come from a model with free or restricted intercepts
<<FA_SCORES_CORR, dependson='FA_SCORES', echo=T, results='markup'>>=
cor(compare.scores[,c(grep("EXT", names(compare.scores), value=T))])
cor(compare.scores[,c(grep("INT", names(compare.scores), value=T))])
@

The gradient of scores by income quintile are very similar
<<FA_SCORES_INEQ, dependson='FA_SCORES', echo=T, results='markup'>>=

# free intercepts
lm(EXTtl ~ factor(incq), data=compare.scores)$coefficients
lm(INTtl ~ factor(incq), data=compare.scores)$coefficients

# constrained intercepts
lm(EXTtli ~ factor(incq), data=compare.scores)$coefficients
lm(INTtli ~ factor(incq), data=compare.scores)$coefficients

# constrained intercepts, common ages
lm(EXTtlir ~ factor(incq), data=compare.scores)$coefficients
lm(INTtlir ~ factor(incq), data=compare.scores)$coefficients

# fill in differences
regs.ext <- list()
regs.int <- list()
for (r in 1:nrow(regcoeffs)) {
  regs.ext[[r]] <- lm(EXT ~ as.factor(incq) + age, data=subset(items.scored[[8]], cohort==regcoeffs[r,1] & sex==regcoeffs[r,2]))
  regcoeffs[r,3] <- round(regs.ext[[r]]$coefficients["as.factor(incq)5"],2)
  regs.int[[r]] <- lm(INT ~ as.factor(incq) + age, data=subset(items.scored[[8]], cohort==regcoeffs[r,1] & sex==regcoeffs[r,2]))
  regcoeffs[r,4] <- round(regs.int[[r]]$coefficients["as.factor(incq)5"],2)
}

% \begin{table}[ht]
% \caption{Estimated means of scores -- non-age adjusted 4-group model} \label{tab:scoremeans}
% \centering
% <<FA_COMP1, dependson='FA_SCORES', results='asis'>>=
% # compare mean factor scores by gender and cohort
% 
% prn <- function(obj) sprintf("%.3f", round(obj,3))
% toaggr <- c("EXT", "INT", "EXTr", "INTr", "EXT.RAW", "INT.RAW", "EXT.RAWr", "INT.RAWr")
% sctable <- aggregate(items.scored[[7]][, toaggr], list(items.scored[[7]]$cohortsex), 
%                    FUN = function (x) prn(mean(x)))
% 
% sctable <- cbind(group = factor(c(1,3,2,4)), sctable)
% levels(sctable$group) <- c("BCS Males", "MCS Males", "BCS Females", "MCS Females")
% sctable <- sctable[order(sctable$group),]
% sctable <- sctable[, !names(sctable) %in% "Group.1"]
% sctable <- as.matrix(sctable,3)
% 
% 
% print.xtable(xtable(sctable,
%                     align="L{0cm}L{2.2cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}",
%                     digits = 3
%                     ),
%              size="\\small",
%              include.rownames=F,
%              include.colnames=F,
%              booktabs = T,
%              add.to.row = list(pos = c(list(0,4)),
%                               command = c(
%                               # header
%                                    "\\toprule
%                                     & \\multicolumn{4}{c}{\\textbf{Scored factors}}
%                                     & \\multicolumn{4}{c}{\\textbf{Scales}} \\\\ \\cmidrule(lr){2-5} \\cmidrule(lr){6-9}
%                                     & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-adjusted} & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-Adjusted} \\\\ \\cmidrule(lr){2-3} \\cmidrule(lr){4-5} \\cmidrule(lr){6-7} \\cmidrule(lr){8-9}
%                       & EXT & INT & EXT & INT & EXT & INT & EXT & INT \\\\",
%                       "\\bottomrule"
%                                     )),
%              hline.after = NULL,  # remove default booktabs lines
%              sanitize.text.function = function(x){x},
%              sanitize.rownames.function = function(x){""}, # to not print row numbers
%              tabular.environment = "tabularx", width = "\\textwidth",
%              floating = FALSE
% )
% 
% @
% \end{table}
% 
% \begin{table}[ht]
% \caption{Estimated means of scores -- age adjusted 4-group model} \label{tab:scoremeans}
% \centering
% <<FA_COMP2, dependson='FA_SCORES', results='asis'>>=
% # compare mean factor scores by gender and cohort
% 
% prn <- function(obj) sprintf("%.3f", round(obj,3))
% toaggr <- c("EXT", "INT", "EXTr", "INTr", "EXT.RAW", "INT.RAW", "EXT.RAWr", "INT.RAWr")
% sctable <- aggregate(items.scored[[8]][, toaggr], list(items.scored[[8]]$cohortsex), 
%                    FUN = function (x) prn(mean(x)))
% 
% sctable <- cbind(group = factor(c(1,3,2,4)), sctable)
% levels(sctable$group) <- c("BCS Males", "MCS Males", "BCS Females", "MCS Females")
% sctable <- sctable[order(sctable$group),]
% sctable <- sctable[, !names(sctable) %in% "Group.1"]
% sctable <- as.matrix(sctable,3)
% 
% 
% print.xtable(xtable(sctable,
%                     align="L{0cm}L{2.2cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}",
%                     digits = 3
%                     ),
%              size="\\small",
%              include.rownames=F,
%              include.colnames=F,
%              booktabs = T,
%              add.to.row = list(pos = c(list(0,4)),
%                               command = c(
%                               # header
%                                    "\\toprule
%                                     & \\multicolumn{4}{c}{\\textbf{Scored factors}}
%                                     & \\multicolumn{4}{c}{\\textbf{Scales}} \\\\ \\cmidrule(lr){2-5} \\cmidrule(lr){6-9}
%                                     & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-adjusted} & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-Adjusted} \\\\ \\cmidrule(lr){2-3} \\cmidrule(lr){4-5} \\cmidrule(lr){6-7} \\cmidrule(lr){8-9}
%                       & EXT & INT & EXT & INT & EXT & INT & EXT & INT \\\\",
%                       "\\bottomrule"
%                                     )),
%              hline.after = NULL,  # remove default booktabs lines
%              sanitize.text.function = function(x){x},
%              sanitize.rownames.function = function(x){""}, # to not print row numbers
%              tabular.environment = "tabularx", width = "\\textwidth",
%              floating = FALSE
% )
% 
% @
% \end{table}
% 
% \begin{table}[ht]
% \caption{Estimated means of scores -- 4-group model, 59-62 months} \label{tab:scoremeans}
% \centering
% <<FA_COMP3, dependson='FA_SCORES', results='asis'>>=
% # compare mean factor scores by gender and cohort
% 
% prn <- function(obj) sprintf("%.3f", round(obj,3))
% toaggr <- c("EXT", "INT", "EXTr", "INTr", "EXT.RAW", "INT.RAW", "EXT.RAWr", "INT.RAWr")
% sctable <- aggregate(items.scored[[9]][, toaggr], list(items.scored[[9]]$cohortsex), 
%                    FUN = function (x) prn(mean(x)))
% 
% sctable <- cbind(group = factor(c(1,3,2,4)), sctable)
% levels(sctable$group) <- c("BCS Males", "MCS Males", "BCS Females", "MCS Females")
% sctable <- sctable[order(sctable$group),]
% sctable <- sctable[, !names(sctable) %in% "Group.1"]
% sctable <- as.matrix(sctable,3)
% 
% 
% print.xtable(xtable(sctable,
%                     align="L{0cm}L{2.2cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}C{1.3cm}",
%                     digits = 3
%                     ),
%              size="\\small",
%              include.rownames=F,
%              include.colnames=F,
%              booktabs = T,
%              add.to.row = list(pos = c(list(0,4)),
%                               command = c(
%                               # header
%                                    "\\toprule
%                                     & \\multicolumn{4}{c}{\\textbf{Scored factors}}
%                                     & \\multicolumn{4}{c}{\\textbf{Scales}} \\\\ \\cmidrule(lr){2-5} \\cmidrule(lr){6-9}
%                                     & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-adjusted} & \\multicolumn{2}{c}{Normal} & \\multicolumn{2}{c}{Age-Adjusted} \\\\ \\cmidrule(lr){2-3} \\cmidrule(lr){4-5} \\cmidrule(lr){6-7} \\cmidrule(lr){8-9}
%                       & EXT & INT & EXT & INT & EXT & INT & EXT & INT \\\\",
%                       "\\bottomrule"
%                                     )),
%              hline.after = NULL,  # remove default booktabs lines
%              sanitize.text.function = function(x){x},
%              sanitize.rownames.function = function(x){""}, # to not print row numbers
%              tabular.environment = "tabularx", width = "\\textwidth",
%              floating = FALSE
% )
% 
% @
% \end{table}

% \clearpage
% \begin{figure}
% \begin{center}
% <<AGEHIST, dev='tikz', fig.height=6, fig.width=6, out.width='.6\\linewidth', cache=TRUE>>=
% @
% \end{center}
% \caption{Distribution of age in the BCS and MCS sample}
% {\scriptsize \emph{Notes}: The figure shows the density of observations at each month of age for each of the two cohorts.  \par}
% \end{figure}
% 
% \clearpage
% \begin{figure}
% <<FACDENS, dev='tikz', fig.height=10, fig.width=10, out.width='1\\linewidth'>>=
% @
% \caption{Density of Scored Factors}
% {\scriptsize \emph{Notes}: The figure shows estimated kernel densities of the scored factors separately by cohort.  \par}
% \end{figure}
% 
% \begin{figure}
% <<FACLOESS_AGE, dev='tikz', fig.height=10, fig.width=10, out.width='1\\linewidth'>>=
% @
% \caption{Scores and child age}
% {\scriptsize \emph{Notes}: The figure plots LOESS of the scored factors against child age. \par}
% \end{figure}
% 
% 
% \begin{figure}
% <<FACRAW, dev='tikz', fig.height=10, fig.width=10, out.width='1\\linewidth'>>=
% @
% \caption{Raw vs Scored Factors}
% {\scriptsize \emph{Notes}: The figure plots the scored factors on the vertical axis against the ``raw" scores on the horizontal axis, i.e. scores obtained by simply summing the number of positive answers to each of the items. \par}
% \end{figure}
% 
% \begin{figure}
% <<FACSCAT, dev='tikz', fig.height=10, fig.width=10, out.width='1\\linewidth'>>=
% @
% \caption{Scatterplots of factor scores by cohort}
% {\scriptsize \emph{Notes}: The figure plots the scored factors separately by cohort. The externalising and internalising scores are plotted against one another. \par}
% \end{figure}
% 
% \begin{figure}
% <<FACINEQ, dev='tikz', fig.height=12, fig.width=10, out.width='1\\linewidth'>>=
% @
% \caption{Distribution of scores across income quintiles}\label{fig:boxscores}
% {\scriptsize \emph{Notes}: The figure plots the distribution of the scored factors across the quintiles in the income distribution. \par}
% \end{figure}
% 
% \begin{figure}
% <<FACLOESS_INC, dev='tikz', fig.height=10, fig.width=10, out.width='1\\linewidth'>>=
% @
% \caption{Scores and family income}
% {\scriptsize \emph{Notes}: The figure plots LOESS of the scored factors against family income. \par}
% \end{figure}

\clearpage
\begin{appendices}

\section{Additional fit comparisons}\label{sec:app_fit}

\begin{table}[ht]
\centering
\small
\caption{Fit indices -- 2-group models}\label{tab:fit2}
<<FA_FIT_RES2, dependson="FA_FIT", results='asis'>>=
require(xtable)

table <- rbind(afitab[[1]],afitab[[2]],afitab[[3]],afitab[[4]],afitab[[5]],afitab[[6]])

print(xtable(table,
                    align="L{0cm}L{.20\\textwidth}C{.03\\textwidth}C{.07\\textwidth}C{.07\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}",
                    digits = 4
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,3,6,9,12,15)),
                              command = c(
                              # header
            "Model & Num. params & $\\chi^2$ & RMSEA & MFI & CFI & G-hat & \\textDelta MFI & \\textDelta CFI & \\textDelta G-hat \\\\ \\midrule
                       \\textbf{(1) No gender split, not age adjusted} \\\\",
                      "\\newline \\textbf{(2) No gender split, age adjusted} \\\\ \\midrule ",
                      "\\newline \\textbf{(3) Males only, not age adjusted} \\\\ \\midrule ",
                      "\\newline \\textbf{(4) Females only, not age adjusted} \\\\ \\midrule ",
                      "\\newline \\textbf{(5) Males only, age adjusted} \\\\ \\midrule ",
                      "\\newline \\textbf{(6) Females only, age adjusted} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

\begin{table}[ht]
\centering
\small
\caption{Fit indices -- 4-group models}\label{tab:fit4}
<<FA_FIT_RES4, dependson="FA_FIT", results='asis'>>=
require(xtable)

table <- rbind(afitab[[7]],afitab[[8]],afitab[[9]],afitab[[11]])

print(xtable(table,
                    align="L{0cm}L{.20\\textwidth}C{.03\\textwidth}C{.07\\textwidth}C{.07\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}C{.06\\textwidth}",
                    digits = 4
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,3,6,9)),
                              command = c(
                              # header
            "Model & Num. params & $\\chi^2$ & RMSEA & MFI & CFI & G-hat & \\textDelta MFI & \\textDelta CFI & \\textDelta G-hat \\\\ \\midrule
                      \\newline \\textbf{(7) not age adjusted} \\\\",
                      "\\newline \\textbf{(8) age adjusted} \\\\\\midrule ",
                      "\\newline \\textbf{(9) 59-62m range} \\\\ \\midrule ",
                      "\\newline \\textbf{(11) MIMIC in age} \\\\ \\midrule "
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)

\clearpage
\section{Latent variable mean structures}\label{sec:app_meancov}

\begin{table}[ht]
\caption{Estimated latent variable means and covariances -- 4 group models} \label{tab:lvpars4}
\centering
<<RES_FA_LV4, dependson="FA_FINAL_PARS", results='asis'>>=
require(xtable)

table <- rbind(lvpars[[7]][[1]],lvpars[[7]][[2]],
               lvpars[[8]][[1]],lvpars[[8]][[2]],
               lvpars[[9]][[1]],lvpars[[9]][[2]]
)

print.xtable(xtable(table,
                    align="L{0cm}L{3cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}",
                    digits = 3
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,2,4,6,8,10,12)),
                              command = c(
                              # header
                                   "\\toprule
                                    & \\multicolumn{4}{c}{\\textbf{BCS}}
                                    & \\multicolumn{4}{c}{\\textbf{MCS}} \\\\ \\cmidrule(lr){2-5} \\cmidrule(lr){6-9}
                                    & Mean & \\multicolumn{2}{c}{Covariance} & Corr. & Mean & \\multicolumn{2}{c}{Covariance} & Corr. \\\\ \\midrule
                      \\textbf{(7) Males (joint, not age adj.)} \\\\ ",
                      "\\newline \\textbf{(7) Females (joint, not age adj.)} \\\\ ",
                      "\\midrule \\newline \\textbf{(8) Males (joint, age adj.)} \\\\ ",
                      "\\newline \\textbf{(8) Females (joint, age adj.)} \\\\ ",
                      "\\midrule \\newline \\textbf{(9) Males (joint, 59-62m)} \\\\ ",
                      "\\newline \\textbf{(9) Females (joint, 59-62m)} \\\\ ",
                      "\\bottomrule"
                                    )),
             hline.after = NULL,  # remove default booktabs lines
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabularx", width = "\\textwidth",
             floating = FALSE
)

\begin{table}[ht]
\caption{Estimated latent variable means and covariances -- 2 group models} \label{tab:lvpars2}
\centering
<<RES_FA_LV2, dependson="FA_FINAL_PARS", results='asis'>>=
require(xtable)

table <- rbind(lvpars[[1]],lvpars[[2]],
               lvpars[[3]],lvpars[[4]],
               lvpars[[4]],lvpars[[6]]
)

print.xtable(xtable(table,
                    align="L{0cm}L{3cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}C{1.2cm}",
                    digits = 3
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,2,4,6,8,10,12)),
                              command = c(
                              # header
                                   "\\toprule
                                    & \\multicolumn{4}{c}{\\textbf{BCS}}
                                    & \\multicolumn{4}{c}{\\textbf{MCS}} \\\\ \\cmidrule(lr){2-5} \\cmidrule(lr){6-9}
                                    & Mean & \\multicolumn{2}{c}{Covariance} & Corr. & Mean & \\multicolumn{2}{c}{Covariance} & Corr. \\\\ \\midrule
                      \\textbf{(1) No gender split, no age adj.} \\\\",
                      "\\textbf{(2) No gender split, age adj.} \\\\",
                      "\\midrule \\newline \\textbf{(3) Males, no age adj.} \\\\ ",
                      "\\newline \\textbf{(4) Females, no age adj.} \\\\ ",
                      "\\midrule \\newline \\textbf{(5) Males, age adj.} \\\\ ",
                      "\\newline \\textbf{(6) Females, age adj.} \\\\ ",
                      "\\bottomrule"
                                    )),
             hline.after = NULL,  # remove default booktabs lines
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabularx", width = "\\textwidth",
             floating = FALSE
)

\clearpage
\section{Estimated model parameters}

\begin{table}[ht]
\caption{Estimated loadings and thresholds -- 2-group models with no age adjustment} \label{tab:allpars2a}
\resizebox{\textwidth}{!}{
<<RES_FA_PAR2A, dependson="FA_FINAL_PARS", results='asis'>>=
require(xtable)

table <- rbind(allpar[[1]],allpar[[3]],allpar[[4]])

@@

                                            & $\\bm{\\lambda}$
                                            & $\\bm{\\tau}_1$
                                            & $\\bm{\\tau}_2$
                                            & $\\bm{\\Var(\\varepsilon)}$ \\\\
                                            Measure & BCS = MCS
                                            & BCS = MCS
                                            & BCS = MCS
                                            & MCS (BCS=1)
                                            & Factor
                                            \\\\ \\midrule
                      \\newline \\textbf{(1) No gender split} \\\\",
                      # additional separators
                      "\\newline \\textbf{(3) Males only} \\\\ \\midrule ",
                      "\\newline \\textbf{(4) Females only} \\\\ \\midrule "
                      )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabularx", width = "\\textwidth",
             floating = FALSE
)

@@

\begin{table}[ht]
\caption{Estimated loadings and thresholds -- 2-group models with age adjustment} \label{tab:allpars2b}
\resizebox{\textwidth}{!}{
<<RES_FA_PAR2B, dependson="FA_FINAL_PARS", results='asis'>>=
require(xtable)

table <- rbind(allpar[[2]],allpar[[5]],allpar[[6]])

@@

                                            & $\\bm{\\lambda}$
                                            & $\\bm{\\tau}_1$
                                            & $\\bm{\\tau}_2$
                                            & $\\bm{\\Var(\\varepsilon)}$ \\\\
                                            Measure & BCS = MCS
                                            & BCS = MCS
                                            & BCS = MCS
                                            & MCS (BCS=1)
                                            & Factor
                                            \\\\ \\midrule
                      \\newline \\textbf{(2) No gender split} \\\\",
                      # additional separators
                      "\\newline \\textbf{(5) Males only} \\\\ \\midrule ",
                      "\\newline \\textbf{(6) Females only} \\\\ \\midrule "
                      )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabularx", width = "\\textwidth",
             floating = FALSE
)

\begin{table}[ht]
\caption{Estimated loadings and thresholds -- 4-group models} \label{tab:allpars4}
\resizebox{\textwidth}{!}{
<<RES_FA_PAR4, dependson="FA_FINAL_PARS", results='asis'>>=
require(xtable)

table <- rbind(allpar[[7]], allpar[[8]], allpar[[9]])

print.xtable(xtable(table,
                    # align = "ll|cccc|cccc"
                    align="L{0cm}L{2.2cm}C{1.8cm}C{1.4cm}C{1.4cm}C{1.7cm}C{1.7cm}C{1.7cm}C{1cm}",
                    digits = 3
                    ),
             size="\\small",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0,11,22)),
                              command = c(
                              # header
                                          " & \\textbf{Loadings}
                                            & \\multicolumn{2}{c}{\\textbf{Thresholds}}
                                            & \\multicolumn{3}{c}{\\textbf{Variances}}
                                            & \\\\ \\cmidrule(lr){2-2} \\cmidrule(lr){3-4} \\cmidrule(lr){5-7}

@@

                                            Measure
                                            & All
                                            & All
                                            & All
                                            & BCS M
                                            & MCS F
                                            & MCS M
                                            & Factor
                                            \\\\ \\midrule 
                                          \\newline \\textbf{(7) not age adjusted} \\\\"
                                          ,
                                          "\\newline \\textbf{(8) age adjusted} \\\\ \\midrule ",
                                          "\\newline \\textbf{(9) 59-62m} \\\\ \\midrule "

                      )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabularx", width = "\\textwidth",
             floating = FALSE
)

@@

\clearpage
\section{Inequality}
